---
title: "Advanced Bioinformatics 2022 assessment"
author: '1810106'
date: "04/04/2022"
output:
---

# TASK 3.1
```{r}
# creating a vector in the provided range of numbers
values = c(5:55)
# summating the values
sum_values = sum(values)
# returning the summed value
print(sum_values)
```

# TASK 3.2
```{r}
# defining a function that sums the input vector
sumfun = function(n) {
  values = c(5:n)
  sum_values = sum(values)
  return(sum_values)
}

# testing function
print(sumfun(10)) # returns 45
print(sumfun(20)) # returns 200
print(sumfun(100)) # returns 5040
```

# TASK 3.3
```{r}
# for loop to generate Fibonacci's sequence

# setting starting values
values = list(0,1)

# calculating Fibonacci values
for (i in c(2:12)) {
  # Fibonacci value = most recent value + previous value
  fib = values[[i]] + values[[i-1]] # must use double square brackets to return object
  # appending Fibonacci value to values list
  values = append(values, fib)
}
# returning the first 12 Fibonacci values
print(values)
```

# TASK 3.4
```{r}
# importing ggplot
library(ggplot2)

# importing mtcars dataset
data(mtcars)

# viewing data
View(mtcars)

# plotting the boxplot using ggplot
boxplot <- ggplot(mtcars, aes(x=gear, y=mpg, fill=factor(gear))) +
  geom_boxplot()
## note: ggplot thinks gears are a continuous variable. Use 'factor' to indicate
## that it is a categorical variable.

# displaying boxplot
boxplot
```

# TASK 3.5
```{r}
# importing the 'cars' dataset
data(cars)

# viewing the data
View(cars)

# fitting the linear model
fit = lm(speed ~ dist, data = cars)

# returning the summary of the linear model
summary(fit)

# information about the cars dataset
?cars
```

PARAMETERS --------------------------------
gradient (slope) = 0.17
intercept = 8.28
standard error for intercept = 0.87438
standard error for distance = 0.01749
units of speed = mph
units of stopping distance = feet

# TASK 3.6
```{r}
# plotting the linear model
lm_plot <- ggplot(cars, aes(x=dist, y=speed)) +
  geom_point() +
  geom_smooth(method='lm')

# displaying the plot
lm_plot
```

# TASK 3.7
```{r}
# importing ggplot
library(ggplot2)

# fitting the linear model
lm_fit = lm(dist ~ I(speed^2), data = cars)
# printing the linear fit
lm_fit

# plotting with ggplot the thinking points and relationship
lm_plot <- ggplot(cars, aes(x=I(speed^2), y=dist)) +
  geom_point() +
  geom_smooth(method='lm')

# displaying plot
lm_plot

# list to store each thinking time
thinking_time_list = list()

# iterate over each row in the cars dataset
## calculate the corresponding thinking time
for (row in 1:nrow(cars)) {
  
  # row varaiables
  speed <- cars[row, "speed"]
  distance <- cars[row, "dist"]
  
  # v = d/t
  ## therefore, t=d/v
  ## d is 8.86, as indicated from the y-intercept from the linear model
  thinking_time = 8.86 / speed
  # appending thinking time to list
  thinking_time_list <- c(thinking_time_list, thinking_time)
}

# converting list to vector
thinking_time_vector = unlist(thinking_time_list)

# adding vector to dataframe
cars$thinking_time <- thinking_time_vector

# view the updated dataframe
View(cars)

```

The fitted linear model indicates that as speed increases, the stopping distance does so as well. The y-intercept of the fit is 8.86, and the gradient is 0.129.

These values were then used to calculate the thinking time.

I believe the output thinking times are reasonable approximations; no negative values are seen,
and assuming the results are in seconds, they indicate a fairly normal reaction time.

However, at slower speeds the thinking time appears to be exceptionally high, such as
a value of 2. A possible explanation for this is that the breaking distance is so short,
it isn't represented very well by the linear model; in other words, the brakes of the car
have an intrinsically different performance at faster speeds compared to slower speeds.

End.